---
title: "Jardim et al. (2023): Supporting Information"
subtitle: "Appendix S1"
author: 
  - Victor L. Jardim
  
crossref:
  fig-prefix: Figure S
  fig-title: Supporting Figure
  subref-labels: alpha a
  thm-prefix: Figure
  thm-title: Figure
  tbl-prefix: Table S
  tbl-title: Supporting Table
format:
  html:
    toc: true
    toc-title: Contents
    toc-location: left
    toc-depth: 5
    toc-expand: true
    theme: minty
    self-contained: true
    embed-resources: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    fig-align: center
    tbl-cap-location: top
    fig-cap-location: bottom
    page-layout: full
    sidebar: true
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console  
execute: 
  cache: refresh
---

```{r}
#| include: false
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, error = TRUE, eval = TRUE, cache = FALSE)
```

This document contains all data analyses presented in the main text as well as supporting information mentioned and extra information on the analysis for those interested (not cross-referenced).

The analysis was organized using the make-line pipeline tool provided by targets [@landau2021]. You will find raw data as well as all functions used in the pipeline in the project's research compendium created with the rcompendium package [@casajusn.2022].

You can take a look at the pipeline in order to have an idea of the different steps for creating the main models and figures. Each target is defined at the \_targets.R file in the research compendium.

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false
network <- targets::tar_visnetwork(reporter = "silent", targets_only = TRUE, exclude = "pal", level_separation = 500, label = c("time", "size"))
```

```{r}
#| echo: false
#| layout: [[-10, 80, -10]]
#| warning: false
#| message: false
#| fig-cap: Analysis pipeline. Total runtime and storage size of each object (target) are shown.
network
```

# Table of figures

| Supporting      | Main text  |
|-----------------|------------|
| @fig-data       | @thm-fig-1 |
| @fig-pcafullsel | @thm-fig-2 |
| @fig-corcomp    | @thm-fig-3 |
| @fig-clust      | @thm-fig-4 |
| @fig-cor        |            |
| @fig-phylord    |            |
| @fig-fam        |            |
| @fig-psN2       |            |
| @fig-psJ        |            |
| @fig-pca        |            |
| @fig-HP         |            |
| @fig-repl       |            |

# Table of tables

[***Unfortunately links only redirect to tabsets open by default but not hidden ones. Please look for tables related to analyses by phylum or by compartment in their respective tabsets.***]{.smallcaps}

| Supporting Tables |
|:-----------------:|
|      @tbl-S1      |
|      @tbl-S2      |
|      @tbl-S3      |
|      @tbl-S4      |
|      @tbl-S5      |
|      @tbl-S6      |
|      @tbl-S7      |
|      @tbl-S8      |
|      @tbl-S9      |
|     @tbl-S10      |
|     @tbl-S11      |
|     @tbl-S12      |
|     @tbl-S13      |
|     @tbl-S14      |

# Packages

```{r}
#| warning: false
#| message: false
library(dplyr)
library(tidyr)
library(vegan)
library(ggplot2)
library(targets)
library(ggrepel)
library(patchwork)
theme_set(theme_light())

```

# Macrofauna data availabity

```{r}
#| layout: [[-10, 80, -10]]
#| tbl-cap: First lines of the original macrofaunal data

tar_load(faunadata)
knitr::kable(head(faunadata))
```

```{r}
#| label: fig-data
#| fig-cap: Number of grab samples taken every year at each point. Empty cells show points where no samples were collected at a given date due to field constraints.
#| message: false

samplespoint <- faunadata %>% 
  mutate(Date=as.factor(format.Date(faunadata$Date,"%Y"))) %>% 
  distinct(Site, Year, Point, Replicate) %>%
  group_by(Site, Year, Point) %>%
  mutate(Year = as.factor(Year)) %>% 
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(State = factor(case_when(count==3 ~ "Good", count < 3 ~ "Sample(s) missing", count > 3 ~ "Too many samples")))

ggplot(samplespoint, aes(x=Year,y=Point,fill=State)) +
  geom_tile(alpha=0.5) +
  scale_fill_manual(values=c("palegreen3","lightblue", "indianred")) +
  geom_text(aes(label=count)) +
  theme(text=element_text(size =15), axis.text.x = element_text(angle = 90, size = 15), legend.position = "bottom")
```

# Habitat Complexity {.unnumbered}

Take a look at the structure of the final complexity dataset

```{r}
#| layout: [[-10, 80, -10]]
#| tbl-cap: First lines of the complexity dataset

tar_load(comp)
knitr::kable(head(comp))
```

Most variables are not normally distributed so we apply a box-cox [@box1964] transformation and check correlations between the transformed variables

```{r}
#| message: false
#| fig-cap: Correlogram of box-cox transformed HC variables
#| fig-width: 14
#| fig-height: 10

tar_read(pairscomp)
```

## PCA of all observations {.unnumbered}

Let's check it without removing highly correlated variables

```{r}
#| fig-width: 20
#| fig-height: 8.75
#| fig-cap: PCA of all observations and all variables

tar_read(pcatot)

```

Let's check the biplots of the selected variables

```{r}
#| fig-width: 20
#| fig-height: 8.75
#| label: fig-pcafullsel
#| fig-cap: PCA of all observations after variable selection

tar_read(pcasel)
```

## PCA of median values

Let's check the biplots of the selected variables

::: {#thm-fig-1}
```{r}
#| fig-width: 14
#| fig-height: 6.125
#| fig-cap: Figure 1 (main text)

tar_read(pcamed)
```
:::

```{r}
#| include: false
#| eval: false
pdf(here::here("figures", "figure-1.pdf"), width = 12, height = 5.25)
tar_read(pcamed)
dev.off()

```

## Correlation between the centroids of the total PCA and the scores of the reduced PCA

```{r}
#| label: fig-corcomp
#| fig-cap: Relation between Full PCA centroids and Reduced PCA scores
#| fig-subcap: 
#|   - "PC1"
#|   - "PC2"
#| layout-ncol: 2

tar_load(corcomp)
tar_load(rv)
plot(corcomp[[1]])
plot(corcomp[[2]])
tar_read(cortestcomp1)
tar_read(cortestcomp2)

```

RV coefficient = **`r rv$rv`**, p-value = **`r rv$p.value`**

## Cluster of median site values

Let's check if we find similar groupings using a ward's cluster

```{r}
#| label: fig-clust
#| fig-width: 10
#| fig-height: 4
#| fig-cap: Ward's cluster of median complexity values at the site level
#| fig-align: left
tar_read(clustcomp)
```

# Check correlations between HC and other environmental variables

```{r}
#| label: fig-cor
#| fig-cap: Correlogram of HC and physical environmental variables
#| fig-width: 14
#| fig-height: 10
tar_read(correl)
```

# Local macrofaunal diversity

```{r}
# Total richness
tar_load(faunaclass)
total_rich <- nrow(faunaclass %>% 
  distinct(., Species)) #725 species

# Total richness by sediment position

epi_rich <- nrow(faunaclass %>% 
  filter(Position == "Epifauna") %>% 
  distinct(., Species)) #341 epifaunal species


inf_rich <- nrow(faunaclass %>% 
  filter(Position == "Infauna") %>% 
  distinct(., Species)) #266  infaunal species


int_rich <- nrow(faunaclass %>% 
  filter(Position == "Interstice") %>% 
  distinct(., Species)) #118 interstitial species
```

```{css, echo = FALSE}
.justify {
  text-align: justify !important
}
```

::: justify
There are **`r total_rich`** species in total, **`r epi_rich`** being epifaunal, **`r inf_rich`** being infaunal, and **`r int_rich`** interstitial.

Taxa from 14 phyla were identified, and Annelida, Arthropoda, Mollusca and Echinodermata were dominant, although Chordata, Cnidaria and Nemertea were also abundant at some sites. Some beds were consistently dominated by polychaetes, while others were always dominated by crustaceans, and a few were more even over the years with dominance shifts between Annelida, Arthropoda and Mollusca (@fig-phylord). Maerl annelid communities were dominated by the mobile predators eunicids (Eunicidae, Lumbrineridae, Dorvilleidae), the crawling surface deposit feeders spionids (Spionidae), the tube dwelling suspension feeding sabellids (Serpullidae), and the subsurface deposit feeding knot worms (Polygordidae; @fig-fam - A). As for arthropods (@fig-fam - B), the decapods were dominant in most beds, namely the filter-feeding porcelain crabs Pisidia longicornis (Linnaeus, 1767; Porcellanidae), which was consistently dominant. The subsurface deposit feeding Corophiidae amphipods and Apseudiidae tanaids, as well as the scavenging Lysianassoidea amphipods were also abundant in most sites. Finally, maerl mollusc communities (@fig-fam - C) were mostly dominated by small grazing and deposit feeding gastropods (Cerithiidae, Rissoidea and Trochidae), by endo and epifaunal filter-feeding pectinid bivalves (namely Veneridae), and by chitons (Leptochitonidae).
:::

```{r}
#| label: fig-phylord
#| fig-cap: Average Relative Abundance of Main Phyla and Orders
#| fig-subcap: 
#|   - Main Phylla
#|   - Main Orders
#| layout-nrow: 2
#| fig-cap-location: top

tar_read(relab_phy)
tar_read(relab_ord) 


```

> (p.s.: it is normal to have NAs for some orders as higher classification for some of the species is ambiguous or unknown.)

```{r}
#| label: fig-fam
#| fig-cap: "Families Average Relative Abundance for Each Main Phyla"
#| fig-subcap: 
#|   - "Arthropoda"
#|   - "Annelida"
#|   - "Mollusca"
#| fig-cap-location: top
#| layout: [[50,50], [-25,50,-25]]

tar_read(relab_arth)
tar_read(relab_ann) 
tar_read(relab_mol)

```

# Effects of HC in local diversity

::: {#thm-fig-2}
```{r}
#| fig-height: 10
#| fig-width: 16
#| layout-nrow: 2
#| layout-ncol: 1

tar_read(figure3) #figure 2-a
tar_read(psdens) #figure 2-b
```

Figure 2 (main text)
:::

## Observed richness

Check if the relationships between HC and species richness are the same for the total macrofauna as well as for the main phyla and the main compartments (different positions in sediment) separately. Main results shown in @thm-fig-2

::: panel-tabset
### Total macrofauna

Now we first forward select environmental variables and then add sites as a random factor to control for pseudo-replication and other site-related variables that were not taken into account.

```{r}
#| fig-cap: OLS linear regression of Total Macrofauna Richness as a function of Rhodolith complexity (Bed complexity is shown in Figure 2)

tar_read(totrichpc1)
```

We check models with quadratic functions of both PC1 and PC2 to test for a unimodal relationship. Unimodal trends are never significant, suggesting an overall linear relationship. A significant interaction is found between PC1^2^ and both PC2 and PC2^2^ in the last model in @tbl-S1. Since there's a very slight difference in AICs and BICs between the last two models, we keep the more parsimonious one.

```{r}
#| label: tbl-S1
#| tbl-cap: Model statistics of the linear and unimodal models of Richness
tar_load(c(modfixrich1, modfixrichpoly1, modfixrich2, modfixrichpoly2, modfixrich12, modfixrichpoly12))

jtools::export_summs(
  modfixrich1, 
  modfixrichpoly1,
  modfixrich2,
  modfixrichpoly2,
  modfixrich12,
  modfixrichpoly12,
  model.names = c("S ~ PC1", "S ~ PC1+PC1^2", "S ~ PC2", "S ~ PC2+PC2^2", "S ~ PC1*PC2","S ~ (PC1+PC1^2)*(PC2+PC2^2)"),
  statistics = c(
    N = "nobs",
    DF = "df.residual",
    AIC = "AIC",
    BIC = "BIC",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
  coefs = c(
    "Rhodolith \n complexity \n (PC1)" = "PC1_score",
    "Rhodolith \n complexity \n (PC1)" = "`poly(PC1_score, 2)`1",
    "PC1^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "Bed complexity (PC2)" = "`poly(PC2_score, 2)`1",
    "PC2^2" = "`poly(PC2_score, 2)`2",
    "PC1:PC2" = "PC1_score:PC2_score",
    "PC1^2:PC2" ="`poly(PC1_score, 2)`2:`poly(PC2_score, 2)`1",
    "PC1:PC2^2" ="`poly(PC1_score, 2)`1:`poly(PC2_score, 2)`2",
    "PC1^2:PC2^2" ="`poly(PC1_score, 2)`2:`poly(PC2_score, 2)`2"
  ),
  scale = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| label: tbl-S2
#| tbl-cap: Model statistics of the final selected models of Richness
tar_load(c(modfixrich12, modfrsel, modrrsel))

jtools::export_summs(
  modfixrich12,
  modfrsel,
  modrrsel,
  model.names = c("S ~ Complexity", "S ~ Complexity + Environment", "S ~ Complexity + Environment + (1|Site)"),
  statistics = c(
    N = "nobs",
    DF = "df.residual",
    AIC = "AIC",
    BIC = "BIC",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
  coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score",
    "Depth",
    "Exposure" = "Fetch_max",
    "Mean bottom temperature" = "T_mean",
    "Organic Matter %" = "OM",
    "Year"
  ),
  scale = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

### By phylum

```{r}
#| label: tbl-S3
#| tbl-cap: Model statistics of richness as a function of HC for each main phyla 

tar_load(c(modrrann, modrrart, modrrmol, modrrepi, modrrinf, modrrint))

jtools::export_summs(
  modrrann,
  modrrart,
  modrrmol,
  model.names = c("Annelida", "Arthropoda", "Mollusca"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
  coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score"
  ),
  scale = TRUE,
  error_pos = "right",
  bold_signif = .05
)


```

```{r}
#| fig-cap: Main phyla richness as a function of HC
#| fig-subcap: 
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#| layout-ncol: 2
tar_read(annrichpc1)
tar_read(annrichpc2) 
tar_read(artrichpc1)
tar_read(artrichpc2)
tar_read(molrichpc1)
tar_read(molrichpc2)
```

### By compartment

```{r}
#| label: tbl-S4
#| tbl-cap: Model statistics of richness as a function of HC for each compartiment
jtools::export_summs(
  modrrepi,
  modrrinf,
  modrrint,
  model.names = c("Epifauna", "Infauna", "Interstitial fauna"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
    coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score"
  ),
  scale = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: Species richness as a function of HC at each faunal compartment
#| fig-subcap: 
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#| layout-ncol: 2
tar_read(epirichpc1)
tar_read(epirichpc2) 
tar_read(infrichpc1)
tar_read(infrichpc2)
tar_read(intrichpc1)
tar_read(intrichpc2)
```
:::

As we can see, the marginal R2 for the models containing only a subset of the macrofauna are quite low. However, although the effects of HC are less evident in these cases, The pattern for most subsets is very similar to the general pattern, with very slight positive effect of rhodolith complexity and a more important positive effect of bed complexity. Nevertheless Rhodolith complexity seems to play a more important role in driving annelid richness than the other main phyla.

## Density of macrofauna

::: panel-tabset
### Total macrofauna

```{r}
#| label: tbl-S5
#| tbl-cap: "Model statistics of macrofaunal density as a function of HC"


tar_load(c(modfixdens12, modfdsel, modrdsel))
jtools::export_summs(
  modfixdens12,
  modfdsel,
  modrdsel,
  model.names = c("Density ~ Complexity", "Density ~ Complexity + Environment", "Density ~ Complexity + Environment + (1|Site)"),
  statistics = c(
    N = "nobs",
    DF = "df.residual",
    AIC = "AIC",
    BIC = "BIC",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
    coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score",
    "Mean current velocity" = "Current_mean",
    "Depth",
    "Exposure" = "Fetch_max",
    "Gravel",
    "Organic Matter %" = "OM",
    "Mean Temperature" = "T_mean",
    "Year"
  ),
  scale = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

### By phylum

```{r}
#| tbl-cap: Model statistics of Density as a function of HC for each main phyla
#| label: tbl-S6

tar_load(c(modfdann, modrdann, modfdart, modrdart, modfdmol, modrdmol, modfdepi, modrdepi, modfdinf, modrdinf, modfdint, modrdint))

jtools::export_summs(
  modfdann,
  modrdann,
  modfdart,
  modrdart,
  modfdmol,
  modrdmol,
  model.names = c("Annelida LM","Annelida LMM", "Arthropoda LM", "Arthropoda LMM", "Mollusca LM", "Mollusca LMM"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score",
    "Mean current velocity" = "Current_mean",
    "Depth",
    "Exposure" = "Fetch_max",
    "Gravel",
    "Organic Matter %" = "OM",
    "Mean Temperature" = "T_mean",
    "Year"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)


```

```{r}
#| fig-cap: Main phyla density as a function of HC
#| fig-subcap: 
#|   - " "
#|   - " "
#|   - " "
#|   - " "
#|   - " "
#|   - " "
#| layout-ncol: 2
tar_read(anndenspc1)
tar_read(anndenspc2) 
tar_read(artdenspc1)
tar_read(artdenspc2)
tar_read(moldenspc1)
tar_read(moldenspc2)
```

### By compartment

```{r}
#| label: tbl-S7
#| tbl-cap: Model statistics of density of macrofauna as a function of HC for each compartiment

jtools::export_summs(
  modfdepi,
  modrdepi,
  modfdinf,
  modrdinf,
  modfdint,
  modrdint,
  model.names = c("Epifauna LM", "Epifauna LMM", "Infauna LM", "Infauna LMM", "Interstitial fauna LM", "Interstitial fauna LMM"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `Marginal R2` = "r.squared.fixed",
    `Conditional R2` = "r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "PC1_score", 
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "PC1_score:PC2_score",
    "Mean current velocity" = "Current_mean",
    "Depth",
    "Exposure" = "Fetch_max",
    "Gravel",
    "Organic Matter %" = "OM",
    "Mud",
    "Mean Temperature" = "T_mean",
    "Temperature Variation (sd)" = "T_sd",
    "Year"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: Density of macrofauna as a function of HC at each compartment
#| fig-subcap: 
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#|   - ""
#| layout-ncol: 2
tar_read(epidenspc1)
tar_read(epidenspc2) 
tar_read(infdenspc1)
tar_read(infdenspc2)
tar_read(intdenspc1)
tar_read(intdenspc2)
```

```{r}
#| fig-cap: Macrofaunal densities as a function of HC for each faunal compartment
#| fig-width: 16
#| fig-height: 10
tar_read(psdtrait)

```
:::

```{r}
#| include: false
#| eval: false
pdf(here::here("figures", "figure-2.pdf"), width = 16, height = 10)
tar_read(psdens)
dev.off()
```

## N2

```{r}
#| label: fig-psN2 
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Regression coefficients of LMs and LMM of Simpson's inverse diversity (N2) as a function of HC and physical environmental variables.
#| 
#| warning: false
#| message: false
tar_read(psN2)
```

## J

```{r}
#| label: fig-psJ
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Regression coefficients of LMs and LMM of Pielou's evenness (J) as a function of HC and physical
#| warning: false
#| message: false
tar_read(psJ)
```

# Regional patterns of biodiversity

## PCA

The first axis of the PCA of the log-chord transformed community data (@fig-pca) highlights a very similar gradient in macrofaunal diversity to that of rhodolith complexity (Figure 1) and accounts for around 14 % of the variance in assemblages. The second axis mostly captures the latitudinal gradient in the study area, which explains around 9.3 % of the total variance. These gradients are mostly explained by differences in the presence and abundance of a few key species (the ones with at least 55 % of their variance represented in the first two axes). The crustacean Athanas nitescens (Leach, 1814) is very rare or never found in most sites except for Rozegat, Keraliou and Trevignon. These sites are also grouped by the presence of Pisidia longicornis, which is found in most sites but is a lot more abundant at the bay of Brest and Trevignon, reaching up to 90 % of relative abundance during some years in the latter. The crustacean Leptocheirus tricristatus (Chevreux, 1887) is present in most sites, with the exception of Rozegat and Keraliou, and is more abundant in Belle-Ile and Meaban (reaching up to 12 % of relative abundance). Finally, the polychaete Glycera lapidum is more abundant in Saint-Brieuc and Camaret but is very rare or absent in sites at the other end of the gradient.

```{r}
#| fig-width: 14
#| label: fig-pca
#| fig-height: 6.125
#| fig-cap: "Principal Component Analysis of log-chord transformed macrofaunal species abundances. The first two axes represent around 23 % of the variance in community composition and structure. Left panel: distance biplot (scaling one) with outer plots representing the density of points for each site. Right panel: correlation biplot (scaling 2) showing species with for which at least 0.55 of their total variance is represented in the ordination plane."

tar_read(faunapca)
```

## Cluster

```{r}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: Ward's cluster of median complexity values at the site level
#| fig-align: left
tar_read(faunaclust)
```

# Effects of HC on regional diversity patterns

## Redundancy analysis

After variable selection, we add sites as a factor to understand if we describe them well with the environmental variables we chose. We can look at the first two axes to have an idea of the relationships between the different explanatory variables.

```{r}
#| label: RDA-triplot
#| fig-cap: RDA triplot showing the variability in species composition related to physical environmental variables and habitat complexity. The first two axes are shown. 
#| fig-height: 8
#| fig-width: 20

tar_read(rdafselcomp)
tar_read(rdafseltemp)
tar_read(rdafselgranulo)
tar_read(rdafselhydro)
tar_load(rdasite)
tar_read(triplotrda)
```

Adjusted R^2^ **\`r RsquareAdj(rdasite)\$adj.r.squared** We check the model's validity. Only 99 permutations were chosen for the validation by axis and by term here as they are really time consuming but the results are the same with 999 permutations. (Output is collapsed, click on code to see)

```{r}
#| collapse: true
tar_read(aovsite)
tar_read(aovaxis)
tar_read(aovterm)
```

## Hierarchical partitioning

::: {#thm-fig-3}
```{r}
#| fig-cap: Figure 3 (main text)
#| fig-width: 16
#| fig-height: 6
tar_read(upsetmain)

```
:::

```{r}
#| include: false
#| eval: false
pdf(here::here("figures", "figure-6.pdf"), width = 16, height = 10)
tar_read(upsetmain)
dev.off()
```

```{r}
#| label: fig-HP
#| fig-cap: Hierarchical partitioning separating Rhodolith and Bed complexity
#| fig-width: 16
#| fig-height: 6
tar_read(upsetsep)
```

# Community stability

::: {#thm-fig-4}
```{r}
#| fig-height: 6
#| fig-width: 12
tar_read(bdplot)
```

```{r}
#| fig-width: 16
#| fig-height: 10
#| fig-subcap: 
#|   - "Effects of HC on community stability by compartiment"
#|   - "Effects of HC on community stability by phylum - not shown in main text"
#| layout-nrow: 2
#| layout-ncol: 1

tar_read(bdtrait)
tar_read(bdphyl)
```
:::

```{r}
#| include: false
#| eval: false
pdf(here::here("figures", "figure-4-A.pdf"), width = 16, height = 10)
tar_read(bdplot)
dev.off()

pdf(here::here("figures", "figure-4-B.pdf"), width = 16, height = 10)
tar_read(bdtrait)
dev.off()
```

```{r}
#| label: fig-repl
#| fig-height: 6
#| fig-width: 12
#| fig-cap: Relative importance of the Replacement component of temporal BD
#| panel: fill
tar_read(replplot)
```

# Effects of HC on community stability

::: panel-tabset
## Total Macrofauna

```{r}
#| label: tbl-S8
#| tbl-cap: Models statistics of temporal diversity metrics as a function of HC
tar_load(c(modbdtot, modrdftot, modrptot))
jtools::export_summs(
  modbdtot,
  modrptot,
  modrdftot,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: Effects of HC on community stability
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdtotpc1)
tar_read(bdtotpc2)
tar_read(rdtotpc1)
tar_read(rdtotpc2)
tar_read(rptotpc1)
tar_read(rptotpc2)
```

## By phylum

::: panel-tabset
### Annelida

```{r}
#| label: tbl-S9
#| tbl-cap:  Models statistics of annelid temporal diversity metrics as a function of HC
tar_load(c(modbdann, modrdfann, modrpann))
jtools::export_summs(
  modbdann,
  modrpann,
  modrdfann,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: Effects of HC on annelid community stability
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdannpc1)
tar_read(bdannpc2)
tar_read(rdannpc1)
tar_read(rdannpc2)
tar_read(rpannpc1)
tar_read(rpannpc2)
```

### Arthropoda

```{r}
#| label: tbl-S10
#| tbl-cap:  Models statistics of arthropod temporal diversity metrics as a function of HC
tar_load(c(modbdart, modrdfart, modrpart))
jtools::export_summs(
  modbdart,
  modrpart,
  modrdfart,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)

```

```{r}
#| fig-cap: "Effects of HC on arthropod community stability"
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdartpc1)
tar_read(bdartpc2)
tar_read(rdartpc1)
tar_read(rdartpc2)
tar_read(rpartpc1)
tar_read(rpartpc2)
```

### Mollusca

```{r}
#| label: tbl-S11
#| tbl-cap:  Models statistics of mollusc temporal diversity metrics as a function of HC

tar_load(c(modbdmol, modrdfmol, modrpmol))
jtools::export_summs(
  modbdmol,
  modrpmol,
  modrdfmol,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: "Effects of HC on mollusc community stability"
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdmolpc1)
tar_read(bdmolpc2)
tar_read(rdmolpc1)
tar_read(rdmolpc2)
tar_read(rpmolpc1)
tar_read(rpmolpc2)
```
:::

## By compartment

::: panel-tabset
### Epifauna

```{r}
#| label: tbl-S12
#| tbl-cap:  Models statistics of epifaunal temporal diversity metrics as a function of HC

tar_load(c(modbdepi, modrdfepi, modrpepi))
jtools::export_summs(
  modbdepi,
  modrpepi,
  modrdfepi,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)

```

```{r}
#| fig-cap: Effects of HC on epifaunal community stability
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdepipc1)
tar_read(bdepipc2)
tar_read(rdepipc1)
tar_read(rdepipc2)
tar_read(rpepipc1)
tar_read(rpepipc2)
```

### Infauna

```{r}
#| label: tbl-S13
#| tbl-cap:  Models statistics of infaunal temporal diversity metrics as a function of HC

tar_load(c(modbdinf, modrdfinf, modrpinf))
jtools::export_summs(
  modbdinf,
  modrpinf,
  modrdfinf,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)

```

```{r}
#| fig-cap: "Effects of HC on infaunal community stability"
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdinfpc1)
tar_read(bdinfpc2)
tar_read(rdinfpc1)
tar_read(rdinfpc2)
tar_read(rpinfpc1)
tar_read(rpinfpc2)
```

### Interstitial fauna

```{r}
#| label: tbl-S14
#| tbl-cap:  Models statistics of interstitial fauna temporal diversity metrics as a function of HC

tar_load(c(modbdint, modrdfint, modrpint))
jtools::export_summs(
  modbdint,
  modrpint,
  modrdfint,
  model.names = c("BDtotal", "Replacement", "Richness Difference"),
  statistics = c(
    N = "nobs",
    `Residual DF` = "df.residual",
    `R2` = "r.squared",
    `Adjusted R2` = "adj.r.squared",
    `F` = "statistic",
    `p-value` = "p.value"
  ),
   coefs = c(
    "Rhodolith complexity (PC1)" = "`poly(PC1_score, 2)`1",
    "Rhodolith complexity (PC1)^2" = "`poly(PC1_score, 2)`2",
    "Bed complexity (PC2)" = "PC2_score",
    "PC1:PC2" = "`poly(PC1_score, 2)`1:PC2_score",
    "PC1ˆ2:PC2" = "`poly(PC1_score, 2)`2:PC2_score"
  ),
  scale = TRUE,
  transform.response = TRUE,
  error_pos = "right",
  bold_signif = .05
)
```

```{r}
#| fig-cap: "Effects of HC on intersitial community stability"
#| layout: [[50,50], [50,50], [50,50]]

tar_read(bdintpc1)
tar_read(bdintpc2)
tar_read(rdintpc1)
tar_read(rdintpc2)
tar_read(rpintpc1)
tar_read(rpintpc2)
```
:::
:::

```{r}
#| include: false
#| eval: false 
pdf(here::here("figures", "figure-8.pdf"), width = 16, height = 10)
tar_read(bdphyl)
tar_read(bdtrait)
dev.off()
```
